---
title: "hw3"
author: "Rhea Pawar"
output: github_document
---

## Problem 0

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(p8105.datasets)
library (ggridges)
library (patchwork)
library (hexbin)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
## Problem 1


#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

## Problem 2

Tidying the accelerometer data

```{r accelerometer, message = FALSE}
acc_df = read_csv("data/accel_data.csv") %>%
  janitor::clean_names() %>% 
  pivot_longer(activity_1:activity_1440,
               names_to = "minute_act",
               names_prefix = "activity_",
               values_to = "physical_activity") %>% 
  mutate(minute_act = as.numeric (minute_act),
         day_type = case_when(day == "MOnday"~"weekday", day == "Tuesday"~"weekday",day == "Wednesday"~"weekday",day == "Thursday"~"weekday",day == "Friday"~"weekday",day == "Saturday"~"weekend", day == "Sunday"~"weekend"))
```

### describing the dataset

`r skimr::skim(acc_df)`

* This dataset contains **`r nrow(acc_df)`** rows(observations) and **`r ncol(acc_df)`** columns(variables). 
* The dataset has variables **`r colnames(acc_df)`**

Aggregating across minutes to find total activity for each day

```{r accelerometer sum, message = FALSE}
sum_acc = acc_df %>%
  select(everything()) %>%
  group_by (day) %>% 
  summarise(total_activity = sum(physical_activity)) %>% 
  pivot_wider(
    names_from = day,
    values_from = total_activity) %>% 
  select( "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday") 

sum_acc %>%   
  knitr::kable()

```
Based on the results it can be seen that the physical activity(in minutes) was higher on Wednesday and Thursday and **highest on Friday**. The activity is **least on Saturday**

```{r plot accelerometer}
plot_acc =acc_df %>% 
  mutate(hour= minute_act/60)
ggplot(plot_acc, aes(x = hour, y = physical_activity, color = day))+
  geom_point(alpha = .5)+geom_line()

```

* From the graph, it can be seen  that the physical activity increases everyday after 5:00 am. It remains almost constant     throughout the day and peaks at 7:00 pm i.e 19:00 hours. Maximum activity can be seen from 7:00 pm to 11: 00 pm i.e 19:00   hrs to 23:00. Least activity is seen from 12:00 am to 5:00 am most likely because individual is sleeping during that time   period.
* The physical activity trend differs by day too. On Sunday maximum activity can be observed from 10 am to 12 pm whereas on   Fridays maximum activity can be observed from 8 pm to 10 pm( 20:00 hrs to 22:00 hrs)

## Problem 3

```{r}
data ("ny_noaa")
ny_noaa = ny_noaa %>% 
  as_tibble(ny_noaa)
```
`r skimr:: skim(ny_noaa)`

